---
title: "portfoliowebsite"
---

```{r}
#| message: false
#| warning: false
#| echo: false
#remove zip code?, parish, telephone number, number of completed surverys footnote,number of completed surverys footnote, additional footnote, state, end date 

#label load packages & import data
library(readr)
library(tidyverse)
library(lubridate)
library(janitor)
library(broom)
library(rsample)
library(yardstick)
library(patchwork)

hosp <- read_csv("HCAHPS-Hospital.csv",na=c("Not Available","Not Applicable"))
                    
```

```{r}
#| message: false
#| warning: false
#remove zip code?, parish, telephone number, number of completed surverys footnote,number of completed surverys footnote, additional footnote, state, end date 

#label load packages & import data
library(readr)
library(tidyverse)
library(lubridate)
here::i_am("Project1.qmd")
hosp <- read_csv(here::here("HCAHPS-Hospital.csv"),na=c("Not Available","Not Applicable"))

```

# Data Wrangling

section explanation

```{r}

#Taking out stuff we don't want
hosp2<-hosp|>
select(!(c("State","Telephone Number","Address","County/Parish","Number of Completed Surveys","Number of Completed Surveys Footnote","Survey Response Rate Percent","Start Date","End Date"))) 
```

```{r}
#patient star rating 
hstar<-hosp2|>
select(!(c("HCAHPS Answer Percent","HCAHPS Linear Mean Value")))|>
 filter(!is.na(`Patient Survey Star Rating`))


#Answer percent

hpercent<-hosp2|>
select(!(c("Patient Survey Star Rating","HCAHPS Linear Mean Value")))|>
  filter(!is.na(`HCAHPS Answer Percent`))
  

#linear mean value
hmean<-hosp2|>
select(!(c("HCAHPS Answer Percent","Patient Survey Star Rating")))|>
  filter(!is.na(`HCAHPS Linear Mean Value`))
```

```{r}
#pivot our HCAHPS COL to get less rows 
hstar<-hstar|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `Patient Survey Star Rating`
  ) 

hpercent<-hpercent|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `HCAHPS Answer Percent`
  ) 

hmean<-hmean|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `HCAHPS Linear Mean Value`
  ) 

 
```

```{r}
library(dplyr)


# Step 2: Convert to numeric safely
hmean$`Overall hospital rating - linear mean score` <- as.numeric(
  as.character(hmean$`Overall hospital rating - linear mean score`)
)


```

```{r}
#Cleaning names

hmeanc<-hmean|>
  mutate(
   Nurse_comm =`Nurse communication - linear mean score`,
   Doctor_comm =`Doctor communication - linear mean score`,
    Staff_resp =`Staff responsiveness - linear mean score`,
   Med_comm =`Communication about medicines - linear mean score`,
   disc_info = `Discharge information - linear mean score`,
   care_trans =`Care transition - linear mean score`,
   cleanliness = `Cleanliness - linear mean score`,
   quietness = `Quietness - linear mean score`,
   overall_rating = `Overall hospital rating - linear mean score`,
   rec_rating = `Recommend hospital - linear mean score`
   
  )

hmeanc$`Nurse communication - linear mean score` <- NULL
hmeanc$`Doctor communication - linear mean score` <- NULL
hmeanc$`Staff responsiveness - linear mean score` <- NULL
hmeanc$`Communication about medicines - linear mean score` <- NULL
hmeanc$`Discharge information - linear mean score` <- NULL
hmeanc$`Care transition - linear mean score` <- NULL
hmeanc$`Cleanliness - linear mean score` <- NULL
hmeanc$`Quietness - linear mean score` <- NULL
hmeanc$`Overall hospital rating - linear mean score` <- NULL
hmeanc$`Recommend hospital - linear mean score` <- NULL
```

# Eda

section explanation

```{r echo=FALSE}
#|echo: false
plot1<- ggplot(data = hmeanc,
       mapping = aes(
         x = Staff_resp,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Staff Responsiveness",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
plot2<-ggplot(data = hmeanc,
       mapping = aes(
         x = Nurse_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Nurse Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot3<-ggplot(data = hmeanc,
       mapping = aes(
         x = Doctor_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Doctor Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot4<-ggplot(data = hmeanc,
       mapping = aes(
         x = Med_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Medicine Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot5<-ggplot(data = hmeanc,
       mapping = aes(
         x = disc_info,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Discharge Information",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
plot6<-ggplot(data = hmeanc,
       mapping = aes(
         x = care_trans,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Care Transition",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot7<-ggplot(data = hmeanc,
       mapping = aes(
         x = cleanliness,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "cleanliness",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot8<-ggplot(data = hmeanc,
       mapping = aes(
         x = quietness,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Medicine Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7)) + theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
```

```{r}
#| message: false
#| warning: false
#| label: show lightbox
#| lightbox: 
#|   match: auto

(plot1+plot2+plot3+plot4+plot5+plot6+plot7+plot8)

```

# Predicitive Model/Model Selection

section explanation

```{r}
set.seed(17)
hmeanc_split <- hmeanc |>
  initial_split(
    prop=0.50
  )  

  hmeanc_training<- training(hmeanc_split)
  hmeanc_test <- testing(hmeanc_split)
```

```{r}
lmall <-lm(
    overall_rating~ Nurse_comm + Doctor_comm + Staff_resp + Med_comm + disc_info + care_trans + cleanliness + quietness,
    data = hmeanc_training
  )

tidy(lmall)
```

```{r}

lmcomm <-glm(
    overall_rating~Nurse_comm + Doctor_comm ,
    data = hmeanc_training 
  )

lmall <-glm(
    overall_rating~Nurse_comm + Doctor_comm + Staff_resp + Med_comm + disc_info + care_trans + cleanliness + quietness  ,
    data = hmeanc_training
  )

lmnull <-glm(
    overall_rating~1 ,
    data = hmeanc_training
  )

tidy(lmcomm)
tidy (lmall)
tidy(lmnull)

hmeanc_pred1<- lmall |>
  augment( newdata = hmeanc_test
  )
head(hmeanc_pred1)

hmeanc_pred1<-hmeanc_pred1|>
  mutate(.resid=overall_rating-.fitted)


ggplot(
  data = hmeanc_pred1,
  mapping= aes(x=.resid)
)+
geom_histogram()

ggplot(data = hmeanc_pred1,
       mapping = aes(
         x = .fitted,
         y = .resid

       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "residual vs fit",
    x = "fitted",
    y = "resdidual")
  
```

```{r}
hmeanc_pred2<- lmcomm |>
  augment( newdata = hmeanc_test)
head(hmeanc_pred2)

hmeanc_pred2<-hmeanc_pred2|>
  mutate(.resid=overall_rating-.fitted)


ggplot(
  data = hmeanc_pred2,
  mapping= aes(x=.resid)
)+
geom_histogram()

ggplot(data = hmeanc_pred2,
       mapping = aes(
         x = .fitted,
         y = .resid

       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "residual vs fit",
    x = "fitted",
    y = "resdidual"
  )
```

```{r}

hmeanc_pred3<- lmnull |>
  augment( newdata = hmeanc_test
  )
head(hmeanc_pred3)

hmeanc_pred3<-hmeanc_pred3|>
  mutate(.resid=overall_rating-.fitted)


ggplot(
  data = hmeanc_pred3,
  mapping= aes(x=.resid)
)+
geom_histogram()

ggplot(data = hmeanc_pred3,
       mapping = aes(
         x = overall_rating,
         y = .resid

       )) +
  geom_point() +
  labs(
    title = "residual vs fit",
    x = "overall rating",
    y = "resdidual"
  )
```

# Project Limitations & Constraints

This subsection should describe potential limitations of the conclusion and/or future work that someone might consider doing to expand on your analysis. Some limitations to discuss might include:

-   What are the real-world consequences of using your data or models? What are the consequences of an incorrect conclusion or prediction?

-   How appropriate was the data you used to answer your question? Consider aspects of validity (variables actually measure what you think they describe) and reliability (variable values would likely be similar if they were measured again).
