---
title: "Hospital Project"
---

```{r}
#| message: false
#| warning: false
#| echo: false
#remove zip code?, parish, telephone number, number of completed surverys footnote,number of completed surverys footnote, additional footnote, state, end date 

#label load packages & import data
library(readr)
library(tidyverse)
library(lubridate)
library(janitor)
library(broom)
library(rsample)
library(yardstick)
library(patchwork)

hosp <- read_csv("HCAHPS-Hospital.csv",na=c("Not Available","Not Applicable"))
                    
```

```{r}
#| message: false
#| warning: false
#| echo: false
#remove zip code?, parish, telephone number, number of completed surverys footnote,number of completed surverys footnote, additional footnote, state, end date 

#label load packages & import data
library(readr)
library(tidyverse)
library(lubridate)
library(dplyr)
here::i_am("Project1.qmd")
hosp <- read_csv(here::here("HCAHPS-Hospital.csv"),na=c("Not Available","Not Applicable"))

```

# Context and Question

The data I looked into in this project is from the Centers for Medicare and Medicaid [Website](https://data.cms.gov/provider-data/dataset/dgck-syfz#data-dictionary). The data set is a listing of various questions and ways of rating hospitals. This inludes star ratings, answers percentages, and mean score which contribute to how well rated patients rate a hospital. Initially what I wanted to explore with this hospital data set was ...

# Data Wrangling

The data set starting off was very messy and hard to work with. To start off looking into this Hospital data set, I took out the unnecessary data for what I was looking into. I used the select function for this, taking out columns in my data set such as State, Address and completion Dates. I then separated out the ways of rating the hospitals into their 3 categories/data sets. With the 3 data sets being linear mean scores out of 100 for specific attributes of the hospital, star ratings out of 5 which are rates similarly to linear mean score and answer percent which tracks what percentages of responses tracked to which specific answers.

```{r}


hosp2<-hosp|>
select(!(c("State","Telephone Number","Address","County/Parish","Number of Completed Surveys","Number of Completed Surveys Footnote","Survey Response Rate Percent","Start Date","End Date"))) 
```

```{r}
#patient star rating 
hstar<-hosp2|>
select(!(c("HCAHPS Answer Percent","HCAHPS Linear Mean Value")))|>
 filter(!is.na(`Patient Survey Star Rating`))


#Answer percent
hpercent<-hosp2|>
select(!(c("Patient Survey Star Rating","HCAHPS Linear Mean Value")))|>
  filter(!is.na(`HCAHPS Answer Percent`))
  

#linear mean value
hmean<-hosp2|>
select(!(c("HCAHPS Answer Percent","Patient Survey Star Rating")))|>
  filter(!is.na(`HCAHPS Linear Mean Value`))
```

## Data Wrangling (Continued)

Separating the rating types wasn't the end of my data wrangling though. My data was still unworkable in its current state. Each thing being rated such as doctor communication or staff responsiveness had its own row. I used the pivot wider function, to widen my data set and give each thing being rated its own column. I then had to as.numeric function to change the data in my linear mean score data set from characters to numbers. Finally I shortened the names in my linear mean value data set, and deleted the old names to make it easier to code and do my eda/model selection.

```{r}
#pivot our HCAHPS COL to get less rows 
hstar<-hstar|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `Patient Survey Star Rating`
  ) 

hpercent<-hpercent|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `HCAHPS Answer Percent`
  ) 

hmean<-hmean|>
  pivot_wider(
    id_cols=`Facility ID`,
    names_from = `HCAHPS Answer Description`,
    values_from = `HCAHPS Linear Mean Value`
  ) 

 
```

```{r}
library(dplyr)


# Step 2: Convert to numeric safely
hmean$`Overall hospital rating - linear mean score` <- as.numeric(
  as.character(hmean$`Overall hospital rating - linear mean score`)
)


```

```{r}
#Cleaning names

hmeanc<-hmean|>
  mutate(
   Nurse_comm =`Nurse communication - linear mean score`,
   Doctor_comm =`Doctor communication - linear mean score`,
    Staff_resp =`Staff responsiveness - linear mean score`,
   Med_comm =`Communication about medicines - linear mean score`,
   disc_info = `Discharge information - linear mean score`,
   care_trans =`Care transition - linear mean score`,
   cleanliness = `Cleanliness - linear mean score`,
   quietness = `Quietness - linear mean score`,
   overall_rating = `Overall hospital rating - linear mean score`,
   rec_rating = `Recommend hospital - linear mean score`
   
  )

hmeanc$`Nurse communication - linear mean score` <- NULL
hmeanc$`Doctor communication - linear mean score` <- NULL
hmeanc$`Staff responsiveness - linear mean score` <- NULL
hmeanc$`Communication about medicines - linear mean score` <- NULL
hmeanc$`Discharge information - linear mean score` <- NULL
hmeanc$`Care transition - linear mean score` <- NULL
hmeanc$`Cleanliness - linear mean score` <- NULL
hmeanc$`Quietness - linear mean score` <- NULL
hmeanc$`Overall hospital rating - linear mean score` <- NULL
hmeanc$`Recommend hospital - linear mean score` <- NULL
```

# Eda

Now that I had my data in the form that I wanted it, I could start on Exploratory analysis to see what is going on. After trying some different type of tables and graphs I settled on making scatter plots. I made a scatter plot for each of the 8 variables in my data that contribute to a hospital's overall linear mean score and mapped them against the overall linear mean score. Each of the graphs showed a positive linear relationship between the variables and the overall linear mean score. This is unsurprising considering the context of the data. If a hospital is good at one area of rating it could be expected that it would be well rated in other areas. Even on a one to one scale it isn't very shocking. The higher a score a hospital gets in one area, the higher overall score it would get.

```{r}
#| label: plot1 chunk
plot1<- ggplot(data = hmeanc,
       mapping = aes(
         x = Staff_resp,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Staff Responsiveness",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
```

This is the code for my first plot. I repeated the same steps for plots 2-8 only changing my y variable and graph name.

```{r}
#| label: plot 2-8 chunk
#| echo: false
plot2<-ggplot(data = hmeanc,
       mapping = aes(
         x = Nurse_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred",se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Nurse Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot3<-ggplot(data = hmeanc,
       mapping = aes(
         x = Doctor_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Doctor Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot4<-ggplot(data = hmeanc,
       mapping = aes(
         x = Med_comm,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Medicine Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot5<-ggplot(data = hmeanc,
       mapping = aes(
         x = disc_info,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Discharge Information",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
plot6<-ggplot(data = hmeanc,
       mapping = aes(
         x = care_trans, 
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Care Transition",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot7<-ggplot(data = hmeanc,
       mapping = aes(
         x = cleanliness,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "cleanliness",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7))+ theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )

plot8<-ggplot(data = hmeanc,
       mapping = aes(
         x = quietness,
         y = overall_rating 
       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "darkblue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Explanatory vs Overall rating",
    x = "Medicine Communication",
    y = "Overall Rating"
  )+theme(plot.title = element_text(size = 7)) + theme(
  axis.title.x = element_text(size = 7),
  axis.title.y = element_text(size = 7) )
```

```{r}
#| message: false
#| warning: false
#| label: show lightbox
#| lightbox: 
#|   match: auto

(plot1+plot2+plot3+plot4+plot5+plot6+plot7+plot8)

```

# Model Selection

section explanation

```{r}
set.seed(17)
hmeanc_split <- hmeanc |>
  initial_split(
    prop=0.50
  )  

  hmeanc_training<- training(hmeanc_split)
  hmeanc_test <- testing(hmeanc_split)
```

```{r}
#| label: big chunk1
lmcomm <-glm(
    overall_rating~Nurse_comm + Doctor_comm ,
    data = hmeanc_training
  )

lmall <-glm(
    overall_rating~Nurse_comm + Doctor_comm + Staff_resp + Med_comm + disc_info + care_trans + cleanliness + quietness  ,
    data = hmeanc_training
  )

lmnull <-glm(
    overall_rating~1 ,
    data = hmeanc_training
  )

tidy(lmcomm)
tidy (lmall)
tidy(lmnull)

hmeanc_pred1<- lmall |>
  augment( newdata = hmeanc_test
  )
head(hmeanc_pred1)

hmeanc_pred1<-hmeanc_pred1|>
  mutate(.resid=overall_rating-.fitted)


ggplot(
  data = hmeanc_pred1,
  mapping= aes(x=.resid)
)+
geom_histogram()

ggplot(data = hmeanc_pred1,
       mapping = aes(
         x = .fitted,
         y = .resid

       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "residual vs fit",
    x = "fitted",
    y = "resdidual")
  
```

```{r}
#| label: big chunk2
hmeanc_pred2<- lmcomm |>
  augment( newdata = hmeanc_test)
head(hmeanc_pred2)

hmeanc_pred2<-hmeanc_pred2|>
  mutate(.resid=overall_rating-.fitted)


ggplot(
  data = hmeanc_pred2,
  mapping= aes(x=.resid)
)+
geom_histogram()

ggplot(data = hmeanc_pred2,
       mapping = aes(
         x = .fitted,
         y = .resid

       )) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", 
              linetype = "dashed", se = FALSE) +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "residual vs fit",
    x = "fitted",
    y = "resdidual"
  )
```

```{r}

hmeanc_pred3<- lmnull |>
  augment( newdata = hmeanc_test
  )
head(hmeanc_pred3)

hmeanc_pred3<-hmeanc_pred3|>
  mutate(.resid=overall_rating-.fitted)



```

```{r}
#| label: bad chunk 1
ggplot(
  data = hmeanc_pred3,
  mapping= aes(x=.resid)
)+
geom_histogram()
```

```{r}
#| label: bad chunk 2
ggplot(data = hmeanc_pred3,
       mapping = aes(
         x = overall_rating,
         y = .resid

       )) +
  geom_point() +
  labs(
    title = "residual vs fit",
    x = "overall rating",
    y = "resdidual"
  ) 
```

```{r}
 hmeanc_pred1 |> 
  mutate(rmse1 = sqrt(mean(.resid^2)))|>
   select(rmse1)

 hmeanc_pred2 |> 
  mutate(rmse2 = sqrt(mean(.resid^2)))|>
   select(rmse2)

```

# Project Limitations & Constraints

This subsection should describe potential limitations of the conclusion and/or future work that someone might consider doing to expand on your analysis. Some limitations to discuss might include:

-   What are the real-world consequences of using your data or models? What are the consequences of an incorrect conclusion or prediction?

-   How appropriate was the data you used to answer your question? Consider aspects of validity (variables actually measure what you think they describe) and reliability (variable values would likely be similar if they were measured again).
